---
title: "Operational Code Analysis (OCA)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Operational Code Analysis (OCA)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This vignette demonstrates Operational Code Analysis (OCA) with leadeR.

## Setup

```{r}
library(leadeR)

spacyr::spacy_initialize()
set.seed(02138)

own_ent <- "United States"
B <- 1000

jfk <- clean_text(jfk19610120)
```

## Computing OCA indices

The `get_oca()` function computes the VICS (Verbs in Context System)
operational code indices from text. It returns the 10 standard indices
(P1--P5 and I1--I5) along with the underlying raw counts.

Set `bootstrap = TRUE` to obtain bootstrap means and variances.

```{r}
res_oca <- get_oca(own_entity = own_ent, text = jfk, bootstrap = TRUE, B = B)
print(res_oca)
```

### Philosophical beliefs (P indices)

| Index | Description                                           |
|-------|-------------------------------------------------------|
| P1    | Nature of the political universe (friendly/hostile)   |
| P2    | Realization of political values (optimism/pessimism)  |
| P3    | Predictability of the political future                |
| P4    | Control over historical development                   |
| P5    | Role of chance                                        |

### Instrumental beliefs (I indices)

| Index | Description                                           |
|-------|-------------------------------------------------------|
| I1    | Strategic approach to goals (cooperative/conflictual) |
| I2    | Intensity of tactics (cooperative/conflictual)        |
| I3    | Risk orientation                                      |
| I4    | Flexibility of tactics (a and b sub-indices)          |
| I5    | Role of chance                                        |
